Identifying `rtrees` Training Data

### 2020-09-29 journals/09-identifying-rtrees-training-data/09-identifying-rtrees-training-data.Rmd
working on condensing journals

>>>>

## Try Old Seeds

The seeds used were selected by Heike based on ones she believes were used at the time of training `rtrees`.

Code for training random forest models based on the random forest fit in Eric's code found [here](https://github.com/erichare/imaging-paper/blob/master/code/full_run.R), which is believed to be the code used to train the original `rtrees`.

```{r}
set.seed(20140501)
rtrees2 <-
  randomForest(
    factor(same_source) ~ ., 
    data = CCFs_withlands %>% select(rtrees_features, same_source), 
    ntree = 300
  )

set.seed(20170222)
rtrees3 <-
  randomForest(
    factor(same_source) ~ ., 
    data = CCFs_withlands %>% select(rtrees_features, same_source), 
    ntree = 300
  )
```

```{r}
set.seed(20170222)
rtrees4 <-
  randomForest(
    factor(same_source) ~ .,
    data = hamby_comparisons_filtered %>% select(rtrees_features, same_source),
    ntree = 300
  )

set.seed(20140501)
rtrees5 <-
  randomForest(
    factor(same_source) ~ .,
    data = hamby_comparisons_filtered %>% select(rtrees_features, same_source),
    ntree = 300
  )
```

```{r}
plot(rtrees2$votes, rtrees3$votes)
plot(rtrees2$votes, bulletxtrctr::rtrees$votes)
plot(rtrees3$votes, bulletxtrctr::rtrees$votes)
```

```{r}
plot(rtrees4$votes, rtrees5$votes)
plot(rtrees4$votes, bulletxtrctr::rtrees$votes)
plot(rtrees5$votes, bulletxtrctr::rtrees$votes)
```

```{r}
plot(sort(rtrees2$votes), sort(bulletxtrctr::rtrees$votes))
plot(sort(rtrees3$votes), sort(bulletxtrctr::rtrees$votes))
plot(sort(rtrees4$votes), sort(bulletxtrctr::rtrees$votes))
plot(sort(rtrees5$votes), sort(bulletxtrctr::rtrees$votes))
```

```{r}
bulletxtrctr::rtrees$confusion
rtrees2$confusion
rtrees3$confusion
rtrees4$confusion
rtrees5$confusion
```

## Distribution Across Seeds

Randomly select 100 seeds (using a random seed):

```{r}
#set.seed(20200917)
set.seed(20200925)
seeds = sample(x = 10000000:100000000, size = 250, replace = FALSE)
seeds
```

Function for fitting random forests to mimic rtrees with different seeds:

```{r}
fit_rf <- function(seed, data) {
  set.seed(seed)
  bullet_rf <-
    randomForest(
      y = factor(data$same_source),
      x = data %>% select(all_of(rtrees_features)),
      ntree = bulletxtrctr::rtrees$ntree,
      mtry = bulletxtrctr::rtrees$mtry
    )
  return(bullet_rf)
}
```

Fit 250 new random forests and save or load:

```{r}
rfs_hamby_comparison_file = "../../../data/rfs_hamby_comparison.rds"
if (!file.exists(rfs_hamby_comparison_file)) {
  rfs_hamby_comparison = map(.x = seeds, .f = fit_rf, data = hamby_comparisons_filtered)
  saveRDS(object = bullet_rfs, file = rfs_hamby_comparison_file)
} else {
  rfs_hamby_comparison = readRDS(rfs_hamby_comparison_file)
}

rfs_CCFs_withlands_file = "../../../data/rfs_CCFs_withlands.rds"
if (!file.exists(rfs_CCFs_withlands_file)) {
  rfs_CCFs_withlands = map(.x = seeds, .f = fit_rf, data = CCFs_withlands)
  saveRDS(object = rfs_CCFs_withlands, file = rfs_CCFs_withlands_file)
} else {
  rfs_CCFs_withlands = readRDS(rfs_CCFs_withlands_file)
}
```

Function for extracting the confusion matrices from the model and put in a nice data frame:

```{r}
extract_confus <- function(model) {
  data.frame(model$confusion) %>%
    rename("FALSE" = "FALSE.", "TRUE" = "TRUE.", "class_error" = "class.error") %>%
    mutate(observed = rownames(model$confusion)) %>%
    pivot_wider(names_from = observed, values_from = everything())  
}
```

Extract the confusion matrices and importance values from all random forest models:

```{r}
hamby_comparison_confus <-
  map_df(.x = hamby_comparison_rfs, .f = extract_confus, .id = "rf_model_id")
CCFs_withlands_confus <-
  map_df(.x = CCFs_withlands_rfs, .f = extract_confus, .id = "rf_model_id")
```

Function for extracting the importance values, computing the rank of the features based on importance, and putting in a data frame:

```{r}
extract_importance <- function(model) {
  data.frame(model$importance) %>%
    mutate(features = rownames(model$importance)) %>%
    arrange(desc(MeanDecreaseGini)) %>%
    mutate(rank = 1:n()) %>%
    rename("importance" = "MeanDecreaseGini") %>%
    select(features, importance, rank)
}
```

```{r}
hamby_comparison_importance <-
  map_df(.x = hamby_comparison_rfs, .f = extract_importance, .id = "rf_model_id")
CCFs_withlands_importance <-
  map_df(.x = CCFs_withlands_rfs, .f = extract_importance, .id = "rf_model_id")
```

### Visualizations of Class Errors

Obtain the class errors from `rtrees` and the paper random forest:

```{r}
rtrees_values <-
  data.frame(
    model = "rtrees",
    class_error_true = rtrees$confusion[2, 3],
    class_error_false = rtrees$confusion[1, 3]
  )
```

Histograms of the class errors (true and false) show that the rtrees values appear to be in the extreme (false) or outside of the new random forest models (true):

```{r fig.width = 10, fig.height = 4.5}
hist_true_hc <-
  hamby_comparison_confus %>%
  ggplot(aes(x = class_error_TRUE)) +
  geom_histogram() +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_true, color = model))

hist_false_hc <- 
  CCFs_withlands_confus %>%
  ggplot(aes(x = class_error_FALSE)) +
  geom_histogram() +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_false, color = model))

hist_true_ccfwl <-
  CCFs_withlands_confus %>%
  ggplot(aes(x = class_error_TRUE)) +
  geom_histogram() +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_true, color = model))

hist_false_ccfwl <- 
  CCFs_withlands_confus %>%
  ggplot(aes(x = class_error_FALSE)) +
  geom_histogram() +
  geom_vline(data = rtrees_values,
             mapping = aes(xintercept = class_error_false, color = model))

plot_grid(hist_true_hc, hist_false_hc, hist_true_ccfwl, hist_false_ccfwl)
```

# LIME Explanation Comparison
